---
title: "Module6-Challenges"
output: html_document
date: "2025-02-09"
---

**Challenge 1**
Loading a data set into a data frame, and summarizing the variables in that data frame using the summary() function. 
```{r}
library(curl)
file <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/Country-Data-2016.csv")
#loading data set by using a github url that links to the raw data
dataframe <- read.csv(file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
#creates a new data frame using the data that was loaded into file. Header = true indicates that there is a header in the data set.
#sep indicates what is used to separate data. Strings as factors converts all the strings to factors. Reminder: factors are 
#strings that have numeric values assigned to them. 
head(dataframe)
#first 6 rows of data
summary(dataframe)
#summary of the different variables
```

**Challenge 2**
Making box plots of the population and area data for the data frame that we already created. Making boxplots again after log transforming the data. Uses boxplot() function and par() function. 
```{r}
par(mfrow = c(2,2))
#organizes the following plots into 2 rows (left number) and 2 columns (right number)

boxplot(dataframe$population, na.rm = TRUE)
boxplot(dataframe$area, na.rm = TRUE)
boxplot(log(dataframe$population), na.rm = TRUE)
boxplot(log(dataframe$area), na.rm = TRUE)
#creates the 4 boxplots
```

**Challenge 3**
Making a histogram of log transformed population and area data. Exploring what happens if freq=TRUE vs if freq=FALSE. Uses hist() function.
```{r}
par(mfrow = c(2,2))
hist(log(dataframe$population), freq = FALSE, col = "red", main = "Plot 1")
hist(log(dataframe$area), freq = FALSE, col = "blue", main = "Plot 2")
#creating two histograms, one for population, one for area. col influences color, main determines the title.

hist(log(dataframe$population), freq = TRUE, col = "green", main = "Plot 3")
hist(log(dataframe$area), freq = TRUE, col = "orange", main = "Plot 4")
#The same histograms, but now freq = TRUE. Frequency = TRUE makes the frequency the y axis. In other words, the count. Frequency = FALSE makes the y axis the proportion 

```

**Challenge 4** 
Make a table to show what the most common form of government is. The table() function makes a table. It can be sorted using the sort() function. 
```{r}
sort(table(dataframe$govt_form), decreasing = TRUE)
#creates a sorted table of government forms, from most to least common.

```

**Challenge 5**
Overlaying boxplots using the ~ operator. 
```{r}
library(curl)
file2 <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(file2, header = TRUE, sep = ",", stringsAsFactors = FALSE)
boxplot(log(d$Body_mass_female_mean)~d$Family)

```

**Challenge 6**
Creating scatterplots using the plot() function.
```{r}
par(mfrow = c(1,2))
plot(x = d$Body_mass_female_mean, y = d$Brain_Size_Female_Mean)
plot(x = log(d$Body_mass_female_mean), y = log(d$Brain_Size_Female_Mean))
#creates two scatterplots, where the x = determines the independent variable, and the y = determines the dependent variable
```

**Challenge 7**
Using ggplot2 to create a scatterplot.
```{r}
library(tidyverse)
plot <- ggplot(data = d, aes(x = log(Body_mass_female_mean), y = log(MaxLongevity_m)))
#ggplot builds plots line by line, notice we do not have to use the $ operator
plot <- plot + geom_point()
##geom_point() makes a scatterplot
plot <- plot +geom_smooth(method = "lm")
#adds a smooth best-fit curve
plot


```

**Challenge 8**
```{r}
d2 <- mutate(d, Binomial = paste(Genus, Species, sep = " "))
#creates new data frame from original data frame, includes new variable, Binomial, which is a concatenation of Genus + Species
d2 <- select(d2, Binomial, Body_mass_female_mean, Body_mass_male_mean, Mass_Dimorphism)
#makes it so that d2 only has 4 variables: binomial, mean female body mass

summary(d2)
head(d2)
#summarizes the variables of d2, and prints the first 6 lines of d2

d2 <- group_by(d2, Binomial)
d2[1:20,]
#organizes the data by binomial, then prints the first 20 lines. Reminder: [row, column] must have a comma in order to determine if you want rows or columns
summarise(d2, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
na.rm = TRUE), avgBMD = mean(Mass_Dimorphism, na.rm = TRUE))
#creates the following summary statistics: the mean male mass among all the primates in the data set, the mean female mass among all the primates in the data set, and the mean sexual dimorphism among all the primates in the data set 

library(tidyverse)
d3 <- filter(d, Family == "Cercopithecidae" | Family == "Atelidae")
#new dataset only including the Cercopithecidae and Atelidae families

plot1 <- ggplot(data = d3, aes(x = Family, y = Body_mass_female_mean))
plot1 <- plot1 + geom_boxplot()
plot1

plot2 <- ggplot(data = d3, aes(x = Family, y = Body_mass_male_mean))
plot2 <- plot2 + geom_boxplot()
plot2

plot3 <- ggplot(data = d3, aes(x = Family, y = Mass_Dimorphism))
plot3 <- plot3 + geom_boxplot()
plot3

```














