---
title: "module-six"
output: html_document
date: "2025-02-04"
---

```{r}
library(curl)
file <-  curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/Country-Data-2016.csv")
dataframe <- read.csv(file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(dataframe)
summary(dataframe$area)
```
**Summary()**
For numeric data: provides summary including minimum, 25th percentile, median, mean, 75th percentile, and maximum of the data, as well as a count of NA (missing values)

For factors: returns a count of a specific string/factor.

**Changing the data type of a column**
use asFactor() or asInteger() or asNumeric() or asLogical()
```{r}
dataframe$govt_form <- as.factor(dataframe$govt_form.) 
summary(dataframe$govt_form)
summary(dataframe$area)
```

**Help Tab**
Can help you figure out what various functions do. 
Located at the top of the bottom right panel.

Can also type ? followed by a topic to get help on it.
Ex: ?integer

**Creating a new column in the dataframe**

```{r}
dataframe$density <- dataframe$population/dataframe$area
summary(dataframe$density)
```

**Showing the 10 most dense countries**
```{r}
dataframe2 <- dataframe[order(-dataframe$density),]
dataframe2[1:10]
```
**Ordering by biggest area**
```{r}
library(curl)
file <-  curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/Country-Data-2016.csv")
dataframe <- read.csv(file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dataframe3 <- dataframe[order(-dataframe$area),]
dataframe3[1:10]
```
**Removing NAs**
na.rm = true

**Boxplot function**
Creates a simple visual of the 5 number summary of numeric data. 
Can only be run on integers or numerics. 
Boxplot()
Reminder: can use str() command to see what columns are what type of data
```{r}
boxplot(as.numeric(dataframe$population))
2.71828^15
```


**Barplots**
barplot().
**The par() command**
This command organizes multiple box plots
Ex: par(mfrow = c(2,3)
The next 6 boxplots will be arranged in a 2 row, 3 column array

```{r}
par(mfrow = c(2, 3))
boxplot(dataframe$population)
boxplot(log(dataframe$population))
boxplot(dataframe$area)
boxplot(log(dataframe$area))
barplot(dataframe$population)
barplot(dataframe$area)
```

**Creating Histograms**
hist() function.
This function will naturally have the frequency (number of observations) on the y axis and the numeric/integer variable on the x axis. 
If you put in the argument freq = FALSE it will instead have the proportion (frequency / number of total observations) of observations on the y axis.
```{r}
hist(log(dataframe$population))
hist(log(dataframe$population), freq = FALSE)
```

**Density plots**
density() creates a density plot
This is essentially a line graph representation of a histogram.

**Note**
For all of these plots, you can use the following commands as arguments.
xlab = "" x axis label
ylab = "" y axis label
more
more
more
more
more

**Tables**
table() can be used to summarize counts and proportions for categorical variables in the dataset.

**~ operator**
Can be used to overlay two different histograms or boxplots

New Data set
```{r}
library(curl)
file2 <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(file2, header = TRUE, sep = ",", stringsAsFactors = FALSE)
boxplot(log(d$Body_mass_female_mean)~d$Family)

```

**Using GGPlot**
A more intuitive and neat way to graph figures. 
To access ggplot, use the command library(ggplot2)
Creates an object which will be the figure
```{r}
library(ggplot2)
p <- ggplot(data = d, aes(x = Family, y = log(Body_mass_female_mean)))  #define the variables
p <- p + geom_boxplot()  #graph them in a boxplot
p <- p + theme(axis.text.x = element_text(angle = 90))  #put x-axis names at 90deg
p <- p + ylab("log(Female Body Mass)")  #rename y-axis title
p  #show me the graph

```

**Scatterplots**
Make a scatterplot with plot(x=XXX, y=YYY)
Plotting relationship between female brain size and female body size, while differentiating families

A pretty scatterplot using ggplot
```{r}

p <- ggplot(data = d, aes(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean),
    color = factor(Family)))  # first, we build a plot object and color points by Family
p <- p + xlab("log(Female Body Mass)") + ylab("log(Female Brain Size)")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())  # then we modify the legend
p  # and, finally, we plot the object

```
We can also add regression lines, using geom_smooth()

**Aggregate statistics and the {dplyr} package**

Calculating summary statistics for groups of obersvations in a data frame

aggregate() function
Within the aggregate function, FUN = "whatever function you need" 

dplyr package includes a set of *verbs* that can perform useful operations on data sets

if there is a function within a package which has the same name as another, you must use the name of the package and a :: before the function so that R knows which function you want. 

*filter* function in dplyr package: can be used to create a new dataframe with certain filters

*arrange* function can organize data by alphabetical (minus sign can be used to reverse order)

*select* function can be used to only select the columns/categories that you want

*rename* function can rename variables, rename(dataframe, new name = old name)

*mutate* function creates a new variable based on already existing variables

*summarize* function 

*group_by()* function

**Piping**
Allows you to run functions sequentially. Takes the result of one function and puts it into the next function. The pipe operator is %>%

**Or operator**
found under the delete key, |

**And operator**
found at the 7 key, &

**Challenge 8**
```{r}
library(ggplot2)
library(curl)
library(dplyr)
file2 <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(file2, header = TRUE, sep = ",", stringsAsFactors = FALSE)

s <- d%>%
    filter(Family =="Cercopithecidae"|Family =="Atelidae")


p <- ggplot(data = s, aes(x = Family, y = Body_mass_female_mean, fill = factor(Family)))# first, we build a plot object and color points by Family
p <- p + xlab("Family") + ylab("Mean Female Body Mass")  # then we modify the axis labels
p <- p + geom_bar(stat='identity')  # then we make a barplot  # then we modify the legend
p  # and, finally, we plot the object






```
