---
title: "Module7-Challenges"
output: html_document
date: "2025-02-09"
---

**Challenge 1**
Given a vector, write a function to determine the geometric mean of the values in a vector.
```{r}
#reminder: to create a function: function name <- function(arguments/inputs){code to run}
geometricMean <- function(x){
  prod(x)^(1/length(x))
  #takes the product of all of the values in the vector x, raises it to the 1/n, which is the same as taking the nth root of it.
}

vector <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 25, 50, 100, 200, 1000)
#creates a vector: a series of data of the same data type
geometricMean(vector)
#runs our function that we just created on the vector we just created

#Note: na.rm = TRUE, and na.omit() can be helpful tools
```

**Challenge 2**
Write a function to calculate the sum of squares for a vector**
```{r}
sos <- function(x){
  sum((x-mean(x))^2)
  #takes the sum of the difference of each value from the mean of the whole vector, squared. 
}
sos(vector)
```

**Challenge 3**
Write a function to calculate the variance of a vector representing a sample.
```{r}
sampVariance <- function(x){
  (sum((x-mean(x))^2))/(length(x)-1)
}
sampVariance(vector)
```

**Challenge 4**
Write a function to calculate standard error
```{r}
stError <- function(x){
  sd(x)/sqrt(length(x))
}
stError(vector)
```

**Normal Distribution Confidence Interval Example**
```{r}
#create a random normal distribution using the rnorm(sample-size, mean, standard deviation) function
#the set.seed() function ensures that each time the function is run, it returns the same set of random numbers
x <- rnorm(10000, 0, 1)

#create a histogram of this norm
hist(x, freq = TRUE)

```

