---
title: "module-six"
output: html_document
date: "2025-02-04"
---

```{r}
library(curl)
file <-  curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/Country-Data-2016.csv")
dataframe <- read.csv(file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(dataframe)
summary(dataframe$area)
```
**Summary()**
For numeric data: provides summary including minimum, 25th percentile, median, mean, 75th percentile, and maximum of the data, as well as a count of NA (missing values)

For factors: returns a count of a specific string/factor.

**Changing the data type of a column**
use asFactor() or asInteger() or asNumeric() or asLogical()
```{r}
dataframe$govt_form <- as.factor(dataframe$govt_form.) 
summary(dataframe$govt_form)
summary(dataframe$area)
```

**Help Tab**
Can help you figure out what various functions do. 
Located at the top of the bottom right panel.

Can also type ? followed by a topic to get help on it.
Ex: ?integer

**Creating a new column in the dataframe**

```{r}
dataframe$density <- dataframe$population/dataframe$area
summary(dataframe$density)
```

**Showing the 10 most dense countries**
```{r}
dataframe2 <- dataframe[order(-dataframe$density),]
dataframe2[1:10]
```
**Ordering by biggest area**
```{r}
library(curl)
file <-  curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/Country-Data-2016.csv")
dataframe <- read.csv(file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dataframe3 <- dataframe[order(-dataframe$area),]
dataframe3[1:10]
```
**Removing NAs**
na.rm = true

**Boxplot function**
Creates a simple visual of the 5 number summary of numeric data. 
Can only be run on integers or numerics. 
Boxplot()
Reminder: can use str() command to see what columns are what type of data
```{r}
boxplot(as.numeric(dataframe$population))
2.71828^15
```


**Barplots**
barplot().
**The par() command**
This command organizes multiple box plots
Ex: par(mfrow = c(2,3)
The next 6 boxplots will be arranged in a 2 row, 3 column array

```{r}
par(mfrow = c(2, 3))
boxplot(dataframe$population)
boxplot(log(dataframe$population))
boxplot(dataframe$area)
boxplot(log(dataframe$area))
barplot(dataframe$population)
barplot(dataframe$area)
```

**Creating Histograms**
hist() function.
This function will naturally have the frequency (number of observations) on the y axis and the numeric/integer variable on the x axis. 
If you put in the argument freq = FALSE it will instead have the proportion (frequency / number of total observations) of observations on the y axis.
```{r}
hist(log(dataframe$population))
hist(log(dataframe$population), freq = FALSE)
```

**Density plots**
density() creates a density plot
This is essentially a line graph representation of a histogram.

**Note**
For all of these plots, you can use the following commands as arguments.
xlab = "" x axis label
ylab = "" y axis label
more
more
more
more
more

**Tables**
table() can be used to summarize counts and proportions for categorical variables in the dataset.

**~ operator**
Can be used to overlay two different histograms or boxplots

New Data set
```{r}
library(curl)
file2 <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(file2, header = TRUE, sep = ",", stringsAsFactors = FALSE)
boxplot(log(d$Body_mass_female_mean)~d$Family)

```







